Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-04-12 paper.qmd
new cubes

The probability of impact for debris object $k$, $p_k$, is expressed as.

@@

where $\rho({\vec{x}})$ is the probability density of an aircraft
and $\sigma(\vec{x})$ is probability density of a debris object in the
position $\vec{x}$.

@@

Debris that reenter the Earth’s atmosphere via drag-induced decay will fly
circular orbits prior to reentry even if initially on an elliptical ones.
In fact drag at perigee will lower apogee and similarly drag at apogee will
lower perigee. But because drag at perigee is higher due to denser atmosphere,
the apogee  will decrease faster than perigee and the orbit will then tend to be
a circular one.
The angular position of the reentry point from the ascending node, 
see @sec-orbitals, is uniformly distributed, $\sigma(\theta) = 1 / (2\pi)$,
due to uncertainties in atmospheric drag and the gravitational effects of
the Equatorial bulge, hence we assume that the impact longitude, $\lambda$,
is uniformely distributed $\sigma(\lambda) = 1/(2\pi)$.

@@

So for example [@patera2008], if a building is located at latitude 30 degrees
north (or south) and has an area of $\qty{1000}{m^2}$ and a piece of debris
reenters the Earth’s atmosphere from an orbit having an inclination of 45 degrees,
the impact probability is given by

@@

$\num{0.046}$

## Aircraft effective exposed area
From Patera

> Even for a ballistic coefficient of 50 lbs/ft2, the flight path angle
> of a reentering object at 40,000 feet is -86.6 degrees, which corresponds
> to nearly vertical descent.
> Since most debris objects have ballistic coefficients equal to or 
> less than 50 lbs/ft2, the vertical descent assumption is a very
> accurate approximation.

The effective expose area, $\varepsilon_i$, is
$$
\varepsilon_i = \frac{s_{a} a_{h} + s_{d} a_{v}}{s_d}
$$
where $s_{a}$ is the average aircraft speed, $s_{d}$ is the average debris 
object fall speed, $a_v$ area of the aircraft as viewed from above,
$a_h$ area of the aircraft as viewed from the from the front of the aircraft.

>>>>

We follow the same mathematical approach of [@patera2008].
The probability, $p_k$, of a debris object $k$, impacting an aircraft over the
area, $a$, is expressed as:

@@

where $\rho({\vec{x}})$ is the probability density of an aircraft in $\vec{x}$,
$\sigma(\vec{x})$ is impact probability density of a debris object in the
position $\vec{x}$ and $\varepsilon_k$ is the effective expose area.

We can compute this probability for each hexagonal cell summing up for every
group of same aircraft type (so with same $\varepsilon_k$).

@@

Debris that reenter the Earth’s atmosphere via drag-induced decay will fly
circular orbits prior to reentry even if initially on elliptical ones.
In fact drag at perigee will lower apogee and similarly drag at apogee will
lower perigee. But because drag at perigee is higher due to denser atmosphere,
the apogee  will decrease faster than perigee and the orbit will then tend to be
a circular one.
The angular position of the reentry point from the ascending node, 
see @sec-orbitals, is uniformly distributed, $\sigma(\theta) = 1 / (2\pi)$,
due to uncertainties in atmospheric drag and the gravitational effects of
the Equatorial bulge, hence we assume that the impact longitude, $\lambda$,
is uniformely distributed $\sigma(\lambda) = 1/(2\pi)$.

@@

So for example if a building is located at latitude 30 degrees
north (or south) and has an area of $\qty{1000}{m^2}$ and a piece of debris
reenters the Earth’s atmosphere from an orbit having an inclination of 45 degrees,
the impact probability is given by

@@

## Aircraft effective exposed area

As from [@patera2008][^balistic_descent], we assume that reentring debris will
descend vertically, which allows to express the effective expose area,
$\varepsilon_i$, as

$$
\varepsilon_i = \frac{s_{a} a_{h} + s_{d_i} a_{v}}{s_{d_i}}
$$

where $s_{a}$ is the average aircraft speed, $s_{d_i}$ is the average fall speed
of the $i$-th debris object, $a_v$ area of the aircraft as viewed from above,
$a_h$ area of the aircraft as viewed from the from the front of the aircraft.

In our calculations we will use

* debris fall speed $s_{d} = \qty{145}{mph}$
* aircraft speed, $s_{a}$, equal to its cruise speed
* aircraft area as viewed from above, $a_v = W \times L$, where $W$ is the
  wingspan, and $L$ the length
* aircraft area as viewed from the front, $a_h = W \times H$, where $W$ is the
  wingspan, and $H$ the height

The values to use have been retrieved from [@performance_database;@skybrary;@doc8643].

[^balistic_descent]: *Even for a ballistic coefficient of 50 lbs/ft2, the flight
path angle of a reentering object at 40,000 feet is -86.6 degrees, which
corresponds to nearly vertical descent.
Since most debris objects have ballistic coefficients equal to or 
less than 50 lbs/ft2, the vertical descent assumption is a very
accurate approximation.*

## Weighting function

::: {#fig-weighting}

```{r weighting}
#| out-width: "100%"
#| echo: false
knitr::include_graphics("media/figures/weighting.png")
```

Weighting function derived from 25 years of reentries. 
:::
